The core transcriptome of mammalian placentas and the divergence of expression with placental shape

### 2016-11-28 manuscript/manuscript.Rnw
fix carter typo; turn of acro macros
LGALS1 is immunomodulating too

%\documentclass[english,12pt]{article}
%\documentclass[5p]{elsarticle_dla}
\documentclass{elsarticle_dla}
\usepackage{fontspec}
\setmainfont{FreeSerif}
\setsansfont{FreeSans}
\setmonofont{FreeMono}
\setmathrm{FreeSerif}
\setmathsf{FreeSans}
\setmathtt{FreeSerif}
%\usepackage[letterpaper,left=1in,right=1in,top=1in,bottom=1in,headheight=14pt]{geometry}
%\usepackage[letterpaper]{geometry}
%\usepackage{url}
\usepackage{graphicx}
%\usepackage{rotating}
%\usepackage{wrapfig}
\usepackage{booktabs}
%\usepackage{float}
\usepackage{multirow}
%\usepackage{setspace}
\usepackage{acro}
\usepackage{array}
\usepackage{longtable}
\usepackage{amsmath}
\usepackage[backend=biber,natbib=true,hyperref=true,citestyle=numeric-comp,style=nature,autocite=inline]{biblatex}
\addbibresource{references.bib}
\usepackage[hyperfigures,bookmarks,colorlinks,linkcolor=black,urlcolor=black,citecolor=blue,breaklinks=true]{hyperref}
% Floats at end for journals
% \usepackage[nolists,figuresfirst,nomarkers,heads]{endfloat}
% \renewcommand{\efloatseparator}{\mbox{}}
%\AtBeginFigures{\newpage \renewcommand{\efloatseparator}{\mbox{}}\section*{Figure Legends}}
%\AtBeginTables{\newpage \section*{Tables}\renewcommand{\efloatseparator}{\mbox{}}}
% \renewcommand{\includegraphics}[2][]{}
\usepackage[nomargin,inline,draft]{fixme}
\newcommand{\DLA}[1]{\textcolor{red}{\fxnote{DLA: #1}}}
\newcommand{\DEW}[1]{\textcolor{purple}{\fxnote{DEW: #1}}}
% \usepackage[x11names,svgnames]{xcolor}
% \usepackage{texshade}
% \usepackage{tikz}
\usepackage{nameref}
\usepackage{xspace}
% this package will stretch the font slightly if necessary.
% \usepackage{microtype}
\usepackage{zref-xr,zref-user}
\zxrsetup{toltxlabel}
\zexternaldocument*{supplemental_material}
\usepackage[capitalize]{cleveref}
\def\UrlBreaks{\do\.\do\@\do\\\do\/\do\!\do\_\do\|\do\;\do\>\do\]%
\do\)\do\,\do\?\do\&\do\'\do+\do\=\do\#%
\do\/%
}%
\input{common_latex}
\tolerance=5000
\emergencystretch=0.1pt

@@

  \textbf{Method} Using RNA-seq, we measured the expression profiles
  of \Sexpr{star.logs[,length(unique(file))]} term placentas of
  \Sexpr{star.logs[,length(unique(species))]} species of mammals, and
  compared the evolution of expression across clades.

  \textbf{Results} We identified a set of
  \Sexpr{core.placenta.transcriptome.long[housekeeping==FALSE,length(unique(human_name))]}
  core genes which are expressed (FPKM $\ge 10$) in all placentas, %
  %
  including genes with immune-modulating properties (\textit{ANXA2},
  \textit{ANXA1}, \textit{S100A11}, and \textit{S100A10}), cell-cell
  interactions (\textit{LAMC1}, \textit{LUM}, and \textit{LGALS1}),
  invasion (\textit{GRB2} and \textit{RALB5}) and trophoblast fusion
  (\textit{ANXA5} and \textit{ANXA1}). %
  We also identified multiple pre-eclampsia associated genes which are
  differentially expressed in \textit{Homo sapiens} when compared to
  the other %
  \Sexpr{star.logs[,length(unique(species))-1]} species. %

@@

The near exclusivity of deep trophoblast invasion in humans has also
led many to speculate on the origin of various disorders, which are
almost exclusively symptomatic of human pregnancy. For example,
pre-eclampsia is a human disease that occurs in 4\% of pregnancies in
the United States
\cite{Chaiworapongsa.ea2014:Pre-eclampsiapart1current}. Pre-eclampsia
can be attributed to the lack of adequate invasion of maternal tissue,
and is noted by the onset of hypertension and proteinuria (or in the
absence of proteinuria, thrombocytopenia, impaired liver function,
renal insufficiency, pulmonary edema, or cerebral or visual
disturbances) %
\cite{Chaiworapongsa.ea2014:Pre-eclampsiapart1current}. Some cases of
pre-eclampsia can develop into eclampsia, a more severe condition
resulting in seizures during pregnancy and death. Lack of invasion in
pre-eclampsia may be attributed to the inefficiency of growth
regulators as well as the inability to sufficiently suppress the
maternal immune response. %
Although there are speculative cases of eclampsia in great apes, 
%\cite{Baird1981:Eclampsiainlowlandgorilla,Stout.ea1969:Glomerularcapillaryendothelialswelling,Thornton.ea1992:Convulsionsinpregnancyin},
it is exceedingly rare in these species
\citeC{arter.ea2011:Evolutionofinvasiveplacentation}.

>>>>

%\documentclass[english,12pt]{article}
%\documentclass[5p]{elsarticle_dla}
\documentclass{elsarticle_dla}
\usepackage{fontspec}
\setmainfont{FreeSerif}
\setsansfont{FreeSans}
\setmonofont{FreeMono}
\setmathrm{FreeSerif}
\setmathsf{FreeSans}
\setmathtt{FreeSerif}
%\usepackage[letterpaper,left=1in,right=1in,top=1in,bottom=1in,headheight=14pt]{geometry}
%\usepackage[letterpaper]{geometry}
%\usepackage{url}
\usepackage{graphicx}
%\usepackage{rotating}
%\usepackage{wrapfig}
\usepackage{booktabs}
%\usepackage{float}
\usepackage{multirow}
%\usepackage{setspace}
\usepackage[macros=false]{acro}
\usepackage{array}
\usepackage{longtable}
\usepackage{amsmath}
\usepackage[backend=biber,natbib=true,hyperref=true,citestyle=numeric-comp,style=nature,autocite=inline]{biblatex}
\addbibresource{references.bib}
\usepackage[hyperfigures,bookmarks,colorlinks,linkcolor=black,urlcolor=black,citecolor=blue,breaklinks=true]{hyperref}
% Floats at end for journals
% \usepackage[nolists,figuresfirst,nomarkers,heads]{endfloat}
% \renewcommand{\efloatseparator}{\mbox{}}
%\AtBeginFigures{\newpage \renewcommand{\efloatseparator}{\mbox{}}\section*{Figure Legends}}
%\AtBeginTables{\newpage \section*{Tables}\renewcommand{\efloatseparator}{\mbox{}}}
% \renewcommand{\includegraphics}[2][]{}
\usepackage[nomargin,inline,draft]{fixme}
\newcommand{\DLA}[1]{\textcolor{red}{\fxnote{DLA: #1}}}
\newcommand{\DEW}[1]{\textcolor{purple}{\fxnote{DEW: #1}}}
% \usepackage[x11names,svgnames]{xcolor}
% \usepackage{texshade}
% \usepackage{tikz}
\usepackage{nameref}
\usepackage{xspace}
% this package will stretch the font slightly if necessary.
% \usepackage{microtype}
\usepackage{zref-xr,zref-user}
\zxrsetup{toltxlabel}
\zexternaldocument*{supplemental_material}
\usepackage[capitalize]{cleveref}
\def\UrlBreaks{\do\.\do\@\do\\\do\/\do\!\do\_\do\|\do\;\do\>\do\]%
\do\)\do\,\do\?\do\&\do\'\do+\do\=\do\#%
\do\/%
}%
\input{common_latex}
\tolerance=5000
\emergencystretch=0.1pt

@@

  \textbf{Method} Using RNA-seq, we measured the expression profiles
  of \Sexpr{star.logs[,length(unique(file))]} term placentas of
  \Sexpr{star.logs[,length(unique(species))]} species of mammals
  representing all major eutherian superordinal clades and marsupials,
  and compared the evolution of expression across clades.

  \textbf{Results} We identified a set of
  \Sexpr{core.placenta.transcriptome.long[housekeeping==FALSE,length(unique(human_name))]}
  core genes which are expressed (FPKM $\ge 10$) in all placentas, %
  %
  including genes with immune-modulating properties (\textit{ANXA2},
  \textit{ANXA1}, \textit{S100A11}, \textit{S100A10}, and \textit{LGALS1}), cell-cell
  interactions (\textit{LAMC1}, \textit{LUM}, and \textit{LGALS1}),
  invasion (\textit{GRB2} and \textit{RALB5}) and syncytialization
  (\textit{ANXA5} and \textit{ANXA1}). %
  We also identified multiple pre-eclampsia associated genes which are
  differentially expressed in \textit{Homo sapiens} when compared to
  the other %
  \Sexpr{star.logs[,length(unique(species))-1]} species. %

@@

The near exclusivity of deep trophoblast invasion in humans has also
led many to speculate on the origin of various disorders, which are
almost exclusively symptomatic of human pregnancy. For example,
pre-eclampsia is a human disease that occurs in 4\% of pregnancies in
the United States
\cite{Chaiworapongsa.ea2014:Pre-eclampsiapart1current}. Pre-eclampsia
can be attributed to the lack of adequate invasion of maternal tissue,
and is noted by the onset of hypertension and proteinuria (or in the
absence of proteinuria, thrombocytopenia, impaired liver function,
renal insufficiency, pulmonary edema, or cerebral or visual
disturbances) %
\cite{Chaiworapongsa.ea2014:Pre-eclampsiapart1current}. Some cases of
pre-eclampsia can develop into eclampsia, a more severe condition
resulting in seizures during pregnancy and death. Lack of invasion in
pre-eclampsia may be attributed to the inefficiency of growth
regulators as well as the inability to sufficiently suppress the
maternal immune response. %
Although there are speculative cases of eclampsia in great apes, 
%\cite{Baird1981:Eclampsiainlowlandgorilla,Stout.ea1969:Glomerularcapillaryendothelialswelling,Thornton.ea1992:Convulsionsinpregnancyin},
it is exceedingly rare in these species
\cite{Carter.ea2011:Evolutionofinvasiveplacentation}.
